---
title: "r exercise 4"
format: html
---

Delete R memory

```{r}
rm(list=ls())

```

# Libraries

```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(sf)
library(terra)
library(tmap)
library(gitcreds)
library(dplyr)
library(SimilarityMeasures)

```

# Preparation

```{r}
testfun <- function(sometext){
  a <- print("hi, ahahahahah!")
  print(sometext)
  return(a)
}

testfun("Hello")

```

Another one

```{r}
#Var1
my_age <- function(birthday){
  now <- Sys.time()
  difftime(now, birthday)
}

#Var 2
my_age1 <- function(birthday){
  difftime(Sys.time(), birthday)
}

#Var 3
my_age2 <- function(birthday, unit){
  difftime(Sys.time(), birthday, unit=unit)
}

my_age("1998-10-30") #Ich bin bereits 8960.41 Tage alt!
my_age("1940-09-17") #Nani ist 30187.45 Tage alt

my_age2("1998-10-30", "mins")

#Var 4
my_age3 <- function(birthday, unit="days"){ #set default unit
  difftime(Sys.time(), birthday)
}

my_age3("1998-10-30", unit="hours")#as well set default unit

```

# Tasks

## Task 1: Write your own functions Import wildboar data

```{r}
ww <- read_csv("wildschwein_BE_2056.csv")
View(ww)

```

Calculate euclidean distances before filtering the data set

```{r}
#mit c(0,1) wir die steplength verändert von ursprünglich direkt Punkt A zu Punkt B, zu vom Punkt A zum übernächsten Punkt. -> Cross-scale movement analysis
euclidean_distance <- function(X = c(0,1), Y = c(0,1)){
  sqrt((lead(X)-X)^2 + (lead(Y)-Y)^2)
}

ww |> 
  mutate(
    stepLength = euclidean_distance(E, N)
  )

```

## Task 2: Prepare Analysis

Finding meet patterns between Rosa and Sabi -\> temporaly and spatially close

```{r}
wwf <- ww |> 
  filter(TierName %in% c("Rosa", "Sabi")) |> 
  #filer(TierName == "Rosa" ¦ TierName == "Sabi") -> makes the same
  filter(DatetimeUTC >= as.POSIXct("2015-04-01 00:00:00", tz="UTC")) |> 
  filter(DatetimeUTC <= as.POSIXct("2015-04-15 23:59:59", tz="UTC"))

#be specific with times, avoid errors


wwf |> 
  ggplot(aes(E,N, colour = TierName))+
  geom_point()+
  coord_equal() #we see if they are at the same place but we need to include the time! 

```

## Task 3: Create Join Key

Include Time -\> join by DatetimeUTC

```{r}
wwf <- wwf |> 
  select(-TierID, -CollarID) |> 
  mutate(
    Datetime_round = lubridate::round_date(DatetimeUTC, "15 minutes") 
  )

```

## Task 4: Measuring distance at concurrent locations

```{r}
wwfr <- wwf |> 
  filter(TierName == "Rosa")

wwfs <- wwf |> 
  filter(TierName == "Sabi")
```

Times are not synchronized, make new colomn that can be used as join key -\> for project: rolling join, allows small time differences

```{r}

rosa_sabi <- dplyr::inner_join(wwfr, wwfs, "Datetime_round", suffix = c("_sabi", "_rosa"))

rosa_sabi <- rosa_sabi |>
  mutate(
    distance = sqrt((E_sabi - E_rosa)^2 + (N_sabi - N_rosa)^2),
    meet = if_else(distance < 100, T, F, missing = NULL)
  ) #|>
  #filter(meet == T)

# -> 11 entries are TRUE for meet

#only the meets

rosab_meet <- rosa_sabi |>
  mutate(
    distance = sqrt((E_sabi - E_rosa)^2 + (N_sabi - N_rosa)^2),
    meet = if_else(distance < 100, T, F, missing = NULL)
  ) |>
  filter(meet)

```

## Task 5: Visualize data

```{r}

as.factor(rosab_meet$meet)

rosab <- rbind(wwfr, wwfs)

tail(rosab)

ggplot()+
  geom_point(rosab, aes(E, N , colour = TierName))+
  geom_point(rosab_meet, aes(E_sabi, N_sabi, shape = meet))+
  coord_equal()

  

ggplot(wwfr, aes(E,N, colour=TierName), alpha = .2)+
  geom_point(wwfs, aes(E,N, colour=TierName), alpha = .2)+
  geom_point(rosab_meet, aes(E_sabi, N_sabi))+
  coord_equal()

#klappt leider noch nicht
# Error in `geom_point()`:
# ! `mapping` must be created by `aes()`
# Backtrace:
#  1. ggplot2::geom_point(wwfs, aes(E, N, colour = TierName), alpha = 0.2)

```

## Task 6 (optional): Visualize data as timecube with plotly

```{r}

```

# Now with the posmo data 

## Task 7: Find "Meet Patterns" in your own tracking data

Load data
```{r}
pp <- read_csv("posmo_1.csv")
View(pp)

```
























































































































